---
title: "Exploratoy Graphs"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(RColorBrewer)
setwd("C:/Users/Nathalie Resoles/Desktop")

# Load the data:
NEI <- readRDS("C:/Users/Nathalie Resoles/Desktop/summarySCC_PM25.rds")
SCC <- readRDS("C:/Users/Nathalie Resoles/Desktop/Source_Classification_Code.rds")
str(NEI)
str(SCC)

```

##Question 1


```{r code q1, echo=FALSE}
totalq1 <- aggregate(Emissions ~ year, NEI, sum)
```

## PLOT 1



```{r emissions, echo=FALSE}
plot(totalq1$year, totalq1$Emissions, type = "o", col = "magenta",
     main = expression("Total US "~ PM[2.5]~ "Emissions by Year"),
     ylab = expression("Total US "~   PM[2.5] ~ "Emissions"), xlab = "Year")
```

#Based on the plot, total emissions have decreased in the US from 1999 to 2008.

##Question 2


```{r baltimore, echo=FALSE}
baltimore <- subset(NEI, NEI$fips == "24510")

totalBaltimore <- aggregate(Emissions ~ year, baltimore, sum)
```

## PLOT 2



```{r emissions baltimore, echo=FALSE}

plot(totalBaltimore$year, totalBaltimore$Emissions, type = "o",
     main = expression("Total Baltimore" ~ PM[2.5] ~ "Emissions by Year"), 
     xlab = "Year", ylab = expression("Total Baltimore "~ PM[2.5] ~ "Emissions"),
     col = "magenta")
```

#Yes, the overall total emissions from PM2.5 decreased in Baltimore from 1999 to 2008.

##Question 3


```{r baltomore year and type, echo=FALSE}
baltimore <- subset(NEI, NEI$fips == "24510")
baltimoretype <- aggregate(Emissions ~ year + type, baltimore, sum)
```

## PLOT 3



```{r baltimore type, echo=FALSE}

ggplot(baltimoretype, aes(year, Emissions, col = type)) +
  geom_line() +
  geom_point() +
  ggtitle(expression("Total Baltimore " ~ PM[2.5] ~ "Emissions by Type and Year")) +
  ylab(expression("Total Baltimore " ~ PM[2.5] ~ "Emissions")) +
  xlab("Year") +
  scale_colour_discrete(name = "Sources") +
  theme(legend.title = element_text(face = "bold"))
```

#All the types of sources such as, non-road, nonpoint, on-road source types have all decreased overall emissions from 1999-2008 in Baltimore City.

#There is a significant increase until 2005 at the "point" source, the point at which it decreases again by 2008 that is just above the starting values.



##Question 4


```{r scc, echo=FALSE}
SCCcoal <- SCC[grepl("coal", SCC$Short.Name, ignore.case = T),]
NEIcoal <- NEI[NEI$SCC %in% SCCcoal$SCC,]
totalCoal <- aggregate(Emissions ~ year + type, NEIcoal, sum)
```

## PLOT 4



```{r coal combustions, echo=FALSE}

ggplot(totalCoal, aes(year, Emissions, col = type)) +
      geom_line() +
      geom_point() +
      ggtitle(expression("Total US" ~ PM[2.5] ~ "Coal Combustion Source Emission by Type and Year")) +
      xlab("Year") +
      ylab(expression("US " ~ PM[2.5] ~ "Coal Emission")) +
      scale_colour_discrete(name = "Sources") +
      theme(legend.title = element_text(face = "bold"))
```

#Emissions from coal combustion related sources of emissions have decreased by about 1/3 from 1999-2008.


##Question 5


```{r using ggplot2, echo= FALSE}
library(ggplot2)
vehicles <- grepl("vehicle", SCC$SCC.Level.Two, ignore.case=TRUE)
vehiclesSCC <- SCC[vehicles,]$SCC
vehiclesNEI <- NEI[NEI$SCC %in% vehiclesSCC,]

baltimorevehiclesNEI <- vehiclesNEI[vehiclesNEI$fips==24510,]
```

## PLOT 5



```{r baltimore vehicle, echo=FALSE}

ggbar<- ggplot(baltimorevehiclesNEI,aes(factor(year),Emissions)) +
  geom_bar(stat="identity",fill="magenta",width=0.75) +
  theme_bw() +  guides(fill=FALSE) +
  labs(x="year", y=expression("Total PM"[2.5]*" Emission (10^5 Tons)")) + 
  labs(title=expression("PM"[2.5]*" Motor Vehicle Source Emissions in Baltimore from 1999-2008"))

print(ggbar)
```

#Emissions from motor vehicle sources have significantly dropped from 1999-2008 in Baltimore City.


##Question 6


```{r type motors, echo= FALSE}

baltLAmotors <- subset(NEI, NEI$fips %in% c("24510","06037") & NEI$type == "ON-ROAD")
baltLAmotorss <- aggregate(Emissions ~ year + fips, baltLAmotors, sum)
```

## PLOT 6



```{r baltimore motors, echo=FALSE}


ggplot(baltLAmotorss, aes(year, Emissions, col = fips)) +
      geom_line() +
      geom_point() +
      ggtitle(expression("Baltimore and Los Angeles" ~ PM[2.5] ~ "Motor Vehicle Emissions by Year")) +
      labs(x = "Year", y = expression(~PM[2.5]~ "Motor Vehicle Emissions") ) +
      scale_colour_discrete(name = "City", labels = c("Los Angeles", "Baltimore")) +
      theme(legend.title = element_text(face = "bold"))
```

#Based on the plot, Los Angeles has seen the greatest changes over time in motor vehicle emissions compared to Baltimore

